---
title: "capstoner"
author: "Lucas Ezequiel Zamora"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Package Description

`capstoner` package reads and plots NOAA data from the [National Centers For Environmental  
Information](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1).


## Functions

Six functions are available; all exported for the user:

Function                 | Description
-------------------------|-------------------------------------------
`eq_clean_data()`        | Clean the NOAA dataset
`eq_clean_location()`    | Clean location from the dataset
`geom_timeline()`        | Geom for plotting a time line of earthquakes
`geom_timeline_label()`  | Labels for the time line plot
`eq_map()`               | Plot a leaflet map of earthquakes 
`eq_create_label()`      | Generates a label for a leaflet map


## Load Packages

```{r}
library(capstoner) 
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
```



## Included data

Data files included within the package are:

```{r}
file_name <- system.file("extdata", package = "capstoner")
data <- readr::read_delim(file_name, quote = "", delim = "\t")
```



## Prepare data for the analysis
The first thing that we must do before starting the analysis with the graphics is to clean the downloaded data since they are incomplete and wrong.

For this we can use the function `eq_clean_data`. The function `eq_clean_location` is used internally by `eq_clean_data`.


```{r, eval=FALSE}
data <- eq_clean_data(data)
```

With the clean data we can already use them to later make the corresponding plots.



## Visualize the NOAA earthquakes dataset on the time
If we would like to visualize the times at which earthquakes occur within certain countries and show the dates on which the earthquakes occur, we can use the `geom_timeline` and `geom_timeline_label`.
The first returns the plot and the second allows the magnitudes to be shown.

```{r, eval=FALSE}
library(dplyr)

data %>%
   capstoner::eq_clean_data() %>%
   dplyr::filter(filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000)
   ) %>%
   ggplot2::ggplot(aes(
       x = date,
       y = country,
       size = eq_primary,
       colour = log(total_deaths)
   )) +
   geom_timeline() +
   geom_timeline_label(n_max = 3)
```



## Visualize the NOAA earthquakes dataset on the space
If we would like to visualize the space at which earthquakes occur within certain countries and show the dates on which the earthquakes occur, we can use the `eq_map` and `eq_create_label`.
The first returns the plot and the second allows to see the information in a friendly format.


```{r, eval=FALSE}
library(dplyr)

data %>%
   capstoner::eq_clean_data() %>%
   dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
   dplyr::mutate(popup_text = eq_create_label(.)) %>%
   eq_map(annot_col = "popup_text")
```